---
title: "sql_freqTable"
author: "RJ"
date: '2022-07-17'
output: html_document
editor_options: 
  chunk_output_type: console
---













SQL Frequency Count


When querying a database,
sometimes you want to see the *frequency count* of different values
appearing in a column / feature, like the Console output 👉


I want to show you 2 useful SQL frequency count techniques:
  1) SQL CASE  (used in Console example)
  2) SQL NTILE

In this video, I am going to use `sqldf::sqldf()`,
which allows us to query datasets using SQL queries. 🚀


Please 🌟PAUSE🌟 the video any time you want to study the code written. ≧◡≦

### First, let me "clean up" IRIS data using `janitor::clean_names()`
```{r}
library(janitor)
iris <- iris |> clean_names()
# janitor::clean_names() was explained in a previous video~

# check iris data
iris |> head(5)
```













### Warm up 🔥
Simple SQL queries to remind us about the famous IRIS dataset
```{r}
# select ALL columns/features to see entire data
sqldf::sqldf("SELECT * FROM iris")

# IRIS is a famous dataset that contains 3 unique species
sqldf::sqldf("SELECT COUNT(DISTINCT species) AS Unique_Species
             FROM iris")

# Frequency count of the 3 unique species: 50 each
sqldf::sqldf("SELECT species AS specie, COUNT(*) AS frequency
             FROM iris
             GROUP BY Species")
```













### SQL CASE
Suppose I want to query something similar to this:

How many samples' `petal_length` are LONGER than x and SHORTER than y, where x < y ?

```{r}
# using R first ...
# number of samples with petal_length 1.9 or shorter
iris$petal_length[iris$petal_length <= 1.9] |> length() # 50
# number of samples whose petal_length longer than 5
iris$petal_length[iris$petal_length > 5]    |> length() # 42


# use SQL CASE to get the same information from database:
sqldf::sqldf("SELECT 
        CASE
          WHEN petal_length <= 1.9 THEN 'short'
          WHEN petal_length BETWEEN 2 AND 5 THEN 'medium'
          ELSE 'long' 
        END as LengthCategory, COUNT(*)
      FROM iris 
      GROUP BY 1
      ORDER BY 1 DESC") 
# short: when petal_length is 1.9 or shorter
# long : when petal_length is longer than 5

# note that "GROUP BY 1" means group by the FIRST column 
# Please PAUSE to understand / read SQL query :) 
```














### SQL NTILE
Suppose I want to create a few intervals with EQUAL number of samples in each of them.

There are 150 samples in IRIS, and I want to create 5 intervals so that 30 samples exist inside each interval (150 = 5 * 30).

We can do it using `SQL NTILE`:
```{r}
# I want 5 intervals so I wrote `NTILE(5)`
sqldf::sqldf("SELECT petal_length, 
              NTILE(5) OVER (ORDER BY petal_length) AS ntile 
              FROM iris")
# notice the `ntile` column:
  # for first  30 samples, ntile is 1
  # for second 30 samples, ntile is 2
  # ...
  # for fifth  30 samples, ntile is 5


# Find the interval boundary petal_lengths like this:
sqldf::sqldf("SELECT ntile, MIN(petal_length), MAX(petal_length)
             FROM (SELECT petal_length, 
                   NTILE(5) OVER (ORDER BY petal_length) AS ntile 
                   FROM iris) 
             GROUP BY ntile")
# for first 30 samples (ntile == 1), min is 1.0 & max is 1.5
# for fifth 30 samples (ntile == 5), min is 5.4 & max is 6.9
```













Alrighty, this is the end of `SQL Frequency Count` video.

Hope you enjoyed it! 
Thank you for watching ◎[▪‿▪]◎











