---
title: "fact_order"
author: "RJ"
date: '2022-06-29'
output: html_document
editor_options: 
  chunk_output_type: console
---







`factor order`


Sometimes it is easier to see the overall pattern
when information is written in some kind of logical "order".

If you see my simple example in Plots panel,
it is easier to see that *C has the highest frequency*, and *H lowest*
with the RIGHT plot than the LEFT plot.

In this short video, I am going to show you how to visualize
**factors in order** 🔥

A lot of explanation I use in this video is from a book titled, 
"R for Data Science" 


Please 🌟PAUSE🌟 the video any time you want to study the code written. ≧◡≦

### First, what package/s are needed?
```{r}
library(tidyverse) # this is ALL we need!
```












### `fct_reorder()`
```{r}
# consider the following made-up data I created
madeup_data <- tibble(
  name  = factor(c("A", "B", "C", "D", "E", "F", "G", "H")),
  freq =         c(32,  12,  48,  23,  20,  19,  37,  8)
)
madeup_data %>% View # let's see the data
madeup_data %>% 
  glimpse() # name is factor ; score is double


# now let's generate a scatter plot, using geom_point()
madeup_data %>% 
  ggplot(aes(x = freq,     # X-Axis
             y = name)) +  # Y-Axis
  geom_point(shape = 15,
             size  = 5) +
  theme_bw() +
  labs(title = "Difficult to interpret the relation of names")


# let's RE-ORDER name using its freq value, using fct_reorder()
madeup_data %>% 
  ggplot(aes(x = freq, 
             y = fct_reorder(name, freq))) +  #RE-ORDER
  geom_point(shape = 15,
             size  = 5) +
  theme_bw() +
  labs(title = "Easier to interpret after RE-ORDER :D")
```







### `fct_infreq()` & `fct_rev()`
```{r}
# consider another made up data:
madeup_data1 <- tibble(
  name  = factor(c("A", "A", "A", "A", "A", 
                   "B", "B", "B",
                   "C", "C", "C", "C", "C", "C", "C",
                   "D")))
madeup_data1 %>% View # let's see the data
madeup_data1 %>% 
  glimpse() # Again, name is factor


# let's generate a bar plot, using geom_bar()
madeup_data1 %>% 
  ggplot(aes(x = name)) + 
  geom_bar() +
  theme_bw() 


# suppose I want to RE-ORDER name according to its frequency
madeup_data1 %>% 
  ggplot(aes(x = fct_infreq(name))) +  # RE-ORDER
  geom_bar() +
  theme_bw() +
  labs(title = "RE-ORDERED :D")


# RE-ORDER in reverse direction compared to above
madeup_data1 %>% 
  ggplot(aes(x = name %>% fct_infreq() %>% 
               fct_rev())) + # reverse order!
  geom_bar() +
  theme_bw() +
  labs(title = "in order :D")
```




Alrighty, this is the end of `factor order` video.

Hope you enjoyed it! 
Thank you for watching ◎[▪‿▪]◎














### Visualization in the beginning....
```{r}
# library(gridExtra)

grid.arrange(a, b, ncol = 2)

a <- madeup_data %>% 
  ggplot(aes(x = freq, 
             y = name)) + 
  geom_point(shape = 15,
             size  = 3) +
  theme_bw() +
  labs(title = "Difficult to interpret",
       subtitle = "LEFT plot")


b <- madeup_data %>% 
  ggplot(aes(x = freq, 
             y = fct_reorder(name, freq))) + 
  geom_point(shape = 15,
             size  = 3) +
  theme_bw() +
  labs(title = "Easy to interpret",
       subtitle = "RIGHT plot")
```




