---
title: "vist"
author: "RJ"
date: '2022-06-24'
output: html_document
editor_options: 
  chunk_output_type: console
---

`vistributions`

If you are studying, reviewing or teaching Statistics, *visualizing/drawing* (statistical) Probability Distributions can sometimes really help you understand\~ (I have an example in my Plots panel)

In this short video, I introduce a useful package `vistributions` that can visualize probability distributions very quickly & easily 😃

If you want to read more about `vistributions`, check out: <https://github.com/rsquaredacademy/vistributions>

Please 🌟PAUSE🌟 the video any time you want to study the code written. ≧◡≦

### First, what packages are needed?

```{r}
library(tidyverse)      # I "always" use tidyverse 🚀
library(vistributions)  # this is the FOCUS 
```

### Normal Distribution

```{r}
# plotting our famous bell 🔔 curve :) 
vistributions::vdist_normal_plot() +
  theme_bw()

# change mean & sd like this
vistributions::vdist_normal_plot(mean = 3, 
                                 sd   = 2) +
  theme_bw()

# 75% lower probability
vistributions::vdist_normal_perc(probs = 0.75,
                                 type = "lower") + 
  theme_bw()

# 75% upper probability
vistributions::vdist_normal_perc(probs = 0.75,
                                 type = "upper") +
  theme_bw()

# 75% 'both' probability
vistributions::vdist_normal_perc(probs = 0.75,
                                 type = "both") +
  theme_bw()

# P(X < 1)
vistributions::vdist_normal_prob(perc = 1) +
  theme_bw()

# P(-1.96 < X < 1.96)
vistributions::vdist_normal_prob(perc = c(-1.96, 1.96)) +
  theme_bw()
```

### Binomial Distribution

```{r}
# Plotting with default n = 10, p = 0.3
vistributions::vdist_binom_plot() +
  theme_bw()

# change n & p 
vistributions::vdist_binom_plot(n = 7,
                                p = 0.2) +
  theme_bw()
# the leftmost bar is about 0.2
# the leftmost bar is probability of failing ALL 7 (= n) trials
# so its calculation would be:
(1 - 0.2)^7 # 0.2097152 , which is indeed about 0.2 ✅
```

### f & chi-square distributions

I am just going to plot 2 distributions together :)

I am using `gridExtra` package.

```{r}
library(gridExtra) # package for drawing 2 distributions together

gridExtra::grid.arrange(vdist_f_plot() + theme_bw(),
                        vdist_chisquare_plot() + theme_bw())
```

Alrighty, this is the end of `vistributions` video.

I encourage you to explore this useful package\~

Hope you enjoyed it! Thank you for watching ◎[▪‿▪]◎
